---
title: "An investigation of the Framingham Heart Study"
author: "By Matthew Curcio"
output: html_document
---


## 1. Experimental Results

```{r LoadData}
options(tinytex.verbose = FALSE)
options(digits = 5) 

#Load Libraries
Libraries <- c("knitr", "readr", "aod")
for (p in Libraries) { 
    library(p, character.only = TRUE)
}

#Load Data
df <- read_csv("CHD_preprocessed.csv", 
     col_types = cols(male = col_factor(levels = c("0", "1")), 
                      education = col_factor(levels = c("0", "1")),
                      currentSmoker = col_skip(), 
                      BPMeds = col_skip(), 
                      prevalentStroke = col_factor(levels = c("0", "1")), 
                      prevalentHyp = col_factor(levels = c("0", "1")), 
                      diabetes = col_factor(levels = c("0", "1")), 
                      TenYearCHD = col_factor(levels = c("0", "1"))))
```


## 2. Logistic Regression Model


```{r cache=TRUE}
mylogit <- glm(TenYearCHD ~ male + age + education + 
                            cigsPerDay + prevalentStroke + prevalentHyp + 
                            diabetes + totChol + sysBP + diaBP + BMI +
                            heartRate + glucose,
               data = df, 
               family = "binomial")

summary(mylogit)
```

- **7 most significant variables**

- Seven predictors have $\small \alpha < 0.05$ (Probabilities). 

- They are significant and associated with acquiring cardiovascular disease.


| Rank | Risk Factor                 |
| :--- | :-------------------------- |
| 1    | Prevalence of Stroke1       |
| 2    | Male1                       |
| 3    | Prevalence of Hypertension1 |
| 4    | Age                         |
| 5    | Cigarettes Per Day          |
| 6    | Systolic Blood Pressure     |
| 7    | Glucose                     |




## 3. Determination of Odds for Two Variables

Odds Given For Male Vs Female

```{r MF, cache=TRUE}
male_test <- with(df, data.frame(male = c("0","1"), #### Factor of Interest ###
                                 age = mean(age),
                                 education = "0",
                                 cigsPerDay = 0,
                                 prevalentHyp = "0",
                                 diabetes = "0",
                                 totChol = mean(totChol),
                                 sysBP = mean(sysBP),
                                 diaBP = mean(diaBP),
                                 BMI = mean(BMI),
                                 heartRate = mean(heartRate),
                                 glucose = mean(glucose),
                                 prevalentStroke = "0"))

# REMEMBER convert male_test from numeric to FACTOR
male_test$male <- as.factor(male_test$male) 

male_test$male <- predict(mylogit, newdata = male_test, type = "response")
```

1. Males: `r male_test$male[[2]]`
2. Female: `r male_test$male[[1]]` 

- **Odds = `r male_test$male[[2]]/male_test$male[[1]]`**



## 4. Odds Given Age


```{r Age, cache=TRUE}
age_test <- with(df, data.frame(male = "0",
                                age = c(20,30,40,50,60,70,80), #### Factor of Interest ###
                                education = "0",
                                cigsPerDay = 0,
                                prevalentHyp = "0",
                                diabetes = "0",
                                totChol = mean(totChol),
                                sysBP = mean(sysBP),
                                diaBP = mean(diaBP),
                                BMI = mean(BMI),
                                heartRate = mean(heartRate),
                                glucose = mean(glucose),
                                prevalentStroke = "0"))

age_test$age <- predict(mylogit, newdata = age_test, type = "response")
```


| Age (years) | Probability Given Age |       Odds Compared to 20 yr old        |
| ----------: | :-------------------: | :-------------------------------------: |
|          20 | `r age_test$age[[1]]` | `r age_test$age[[1]]/age_test$age[[1]]` |
|          30 | `r age_test$age[[2]]` | `r age_test$age[[2]]/age_test$age[[1]]` |
|          40 | `r age_test$age[[3]]` | `r age_test$age[[3]]/age_test$age[[1]]` |
|          50 | `r age_test$age[[4]]` | `r age_test$age[[4]]/age_test$age[[1]]` |
|          60 | `r age_test$age[[5]]` | `r age_test$age[[5]]/age_test$age[[1]]` |
|          70 | `r age_test$age[[6]]` | `r age_test$age[[6]]/age_test$age[[1]]` |
|          80 | `r age_test$age[[7]]` | `r age_test$age[[7]]/age_test$age[[1]]` |


## Notes

For interpretation help:

- https://www.ibm.com/topics/logistic-regression

- https://stats.idre.ucla.edu/other/mult-pkg/faq/general/faq-how-do-i-interpret-odds-ratios-in-logistic-regression/

For analysis help:

- https://stats.idre.ucla.edu/r/dae/logit-regression/

- https://stats.oarc.ucla.edu/r/dae/logit-regression/

- https://bookdown.org/mike/data_analysis/wald-test.html

- https://bookdown.org/mike/data_analysis/hypothesis-testing.html#wald-test

